Historic collections as a tool for assessing the global pollination crisis

### 2018-09-11 revision/SupMat3.Rmd
I think this updates sup mat to match the ms.

---
title: "Historic collections as a tool for assessing the global pollination crisis"
author: "Bartomeus, I.^1^, Stavert, J.R.^2^, Ward, D.^2,3^, and Aguado, O.^4^"
csl: ../data/philb_style.xml
output:
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
bibliography: ../data/references2.xml
---

@@

For illustrative purposes, we mapped the findings of these studies (Figure 1) to show the large differences in bee species richness worldwide, with bee diversity hotspots in mediterranean countries, against the paucity of countries for which we have any local or regional data on bee, syrphid or butterfly declines (raw data: Sup Mat 1). Except for outside of Europe and the US and for non-insect taxa, there are very few or no studies on pollinator declines using historical records. However, there are species-specific examples of historical losses from different parts of the world (e.g., *Bombus dalbhomi*; @Morales2013).

```{r map, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(rworldmap)
map <- read.csv("../data/bee_richness.csv")
size <- read.csv("../data/country_area.csv")
colnames(size) <- c("country", "size")
map0 <- merge(map, size, all.x = TRUE)
#remove some countries with doubtful sampling
#map0 <- subset(map0, !country %in% c("Kuwait", "Guinea-Bissau", "Chad", "Mali", "Mauritania"))
#map0[which(is.na(map0$size)),] #some can be rescued?
#scatter.smooth(log(map0$species) ~ log(map0$size+1))
m <- lm(log(map0$species) ~ log(map0$size+1), na.action = na.exclude) 
map0$res <- residuals(m)
#subset(map0, res < -2)
map2 <- joinCountryData2Map(map0, joinCode="NAME", nameJoinColumn = "country",
                            nameCountryColumn = "distribution")
#need to use ISO's to make sure we don't carry over NA's due to country misspellings
#have plotted countries with missing values as grey (shows we have richness for pretty much everywhere)
par(mfrow = c(2,2))
map3 <- mapCountryData(map2, nameColumnToPlot = "res", missingCountryCol = "white",
               mapTitle = "Bee species richness", numCats = 25, catMethod = "pretty", addLegend = FALSE)
#add a modified legend using the same initial parameters as mapCountryData
do.call(addMapLegend, c(map3, legendLabels="all", legendWidth=0.5))

#need to add all the pannels.
library(plyr)
library(dplyr)
library(reshape2)

@@

map.bee <- joinCountryData2Map(declines.bee.2, joinCode="NAME", nameJoinColumn = "country",
                                   nameCountryColumn = "distribution")

#average richness change
colorss <- c(heat.colors(9))
map.bee.ave <- mapCountryData(map.bee, nameColumnToPlot = "average", colourPalette= colorss,
                                  mapTitle = "Average bee richness change", catMethod = "pretty", addLegend = F)
do.call(addMapLegend, c(map.bee.ave, legendLabels="all", legendWidth=0.5))

@@

map.syrphid <- joinCountryData2Map(declines.syrphid.2, joinCode="NAME", nameJoinColumn = "country",
                            nameCountryColumn = "distribution")

#average richness change
colorss <- c(heat.colors(5), c("#5e977c", "#3d88a5", "#1c79cf"))
map.syrphid.ave <- mapCountryData(map.syrphid, nameColumnToPlot = "average", colourPalette= colorss,
                       mapTitle = "Average syrphid richness change", catMethod = "pretty", addLegend = F)
do.call(addMapLegend, c(map.syrphid.ave, legendLabels="all", legendWidth=0.5))

@@

map.butterfly <- joinCountryData2Map(declines.butterfly.2, joinCode="NAME", nameJoinColumn = "country",
                                     nameCountryColumn = "distribution")

#average richness change
colorss <- c(heat.colors(6))
map.butterfly.ave <- mapCountryData(map.butterfly, nameColumnToPlot = "average", colourPalette=colorss,
                                    mapTitle = "Average butterfly richness change", catMethod = "pretty", addLegend = F)
do.call(addMapLegend, c(map.butterfly.ave, legendLabels="all", legendWidth=0.5))

par(mfrow = c(1,1))
```

**Figure 1:** Global map showing a) bee species richness per area (Data from www.discoverlife.org) calculated as the residuals of the log-log regression between bee species richness per country and country size. This correction accounts for the species-area relationship. Warmer colours indicate higher bee diversity. Note that some African countries may have incomplete listed faunas and that Alaska is included with US values. Countries with available historical changes in (b) bee, (c) syrphid and (d) butterfly richness within the last 100 years. Warmer colours indicate steeper average declines. Countries without data are coloured in white.

@@

#table?
#lhf[which(log(lhf$post1980) > 7.5 & log(lhf$pre1980) < 7.5),]
#lhf[which(log(lhf$post1980) > 7.5 & log(lhf$pre1980) > 7.5),]
#nrow(lhf[which(log(lhf$post1980) < 6.9 & log(lhf$pre1980) < 6.9),])
# poland switcherland need old records! > 10000 vs 300
# sri lanka and french guayana, 
# nicaragua, venezuela, Ecuador..>2000 vs >500.
# Colombia & Panama, zimbawe > 3000 vs >1500
#lhf[which(lhf$country == "ES" | lhf$country == "NZ"),]
col <- ifelse(lhf$country == "ES" | lhf$country == "NZ", "red", "black")
col <- ifelse(log(lhf$post1980) > 7.5 & log(lhf$pre1980) > 7.5, col, "grey")
pchs <- ifelse(lhf$country == "ES" | lhf$country == "NZ", 19, 1)
plot(log(lhf$post1980+1), log(lhf$pre1980+1), xlab = "post1980", ylab = "pre1980", col = col, pch = pchs, las = 1)
temp <- subset(lhf, country %in% c("ES", "NZ"))
text(x = temp$log_post1980, y = temp$log_pre1980+2 , labels = temp$country, col = "red")
abline(h = 7.5, v = 7.5)
#36 countries!
#exp(7.5) #> 1800 records in both axes!
```

@@

boxplot(log(data2$pre1980), log(data2$post1980), names = c("pre1980", "post1980"), ylab = "log rarefied species richness", 
            las = 1, border = "light grey")
points(x = c(rep(1,nrow(data2)), rep (2,nrow(data2))), y = c(log(data2$pre1980), log(data2$post1980)), pch = 1)
for(i in 1:nrow(data2)){
    lines(x = 1:2, y = c(log(data2$pre1980[i]), log(data2$post1980[i])), 
          col = ifelse((data2$pre1980[i] - data2$post1980[i]) > 1, "red", "blue"))
}

#map
data2$change <- ((data2$post1980 - data2$pre1980)*100 / data2$pre1980)
#mean(data2$change)
map.gbif <- joinCountryData2Map(data2, joinCode="ISO2", nameJoinColumn = "countries",
                            nameCountryColumn = "countries")

#average richness change
#colorss <- c("firebrick", "red", "orangered", "orange", "yellow", "wheat", "green", "blue", "darkblue")
colorss <- c(heat.colors(5), c("#5e977c", "#3d88a5", "#1c79cf"))
map.gbif.plot <- mapCountryData(map.gbif, nameColumnToPlot = "change", colourPalette= colorss,
                       mapTitle = "Average bee richness change (%)", catMethod = "pretty", addLegend = F)
do.call(addMapLegend, c(map.gbif.plot, legendLabels="all", legendWidth=0.5))

@@

In contrast to Spain, New Zealand is an isolated oceanic archipelago, with a distinctive pollinator biota and a unique history of human occupation. Much of New Zealand’s pollinator fauna is also relatively depauperate. For example, New Zealand has only 28 native bee species [@Donovan2007], which is a fraction of nearby Australia’s c. 1600 species [@Ascher2015]. However, New Zealand has a surprisingly high diversity of flies (Diptera), which are important pollinators in many ecosystems [@Newstrom2005]. Thus, New Zealand provides a unique system to study long-term changes in pollinator communities, and is unlike continental Europe and the US, which have been the focus of an overwhelming majority of pollinator decline studies.

@@

}
    plot(time,t(xx2), xlab = "", las = 2, ylab = "Number of native bee species", main = "", xaxt = "n", ylim = c(min(xx2, na.rm = TRUE)- max(se), max(xx2, na.rm = TRUE)+ max(se)))
    axis(1, at = time ,labels = lab, las = 2, cex.axis = 0.7)
    arrows(time,as.numeric(xx2)+as.numeric(se), time, as.numeric(xx2)-as.numeric(se), angle=90, length = 0)
    abline(m, lty = 2)

#############################################
#Resampling Analysis for exotic species 
#############################################

#length(which(exotic$year < 1900))
exotic <- subset(exotic, year > 1800)
for (i in 9){ #error saying break are not unique - how to fix?
    break1 <- quantile(exotic$year, probs = seq(0,1,1/6), na.rm = TRUE)
    x <- c(1916,1952,1953,1969,1971,1975,1978,1983,1991,2011)
    lab <- c(1:i)
    for(j in 1:i){
        lab[j] <- paste(as.numeric(x)[j],as.numeric(x)[j+1], sep = "-")
    }
    exotic$time_period_quant  <- cut(exotic$year, breaks = x, labels = lab)
    comm2 <- table(exotic$time_period_quant, factor(exotic$Gen_sp))
    (x2 <- rarefy(comm2,25, se = TRUE)) #same result as my resampling...
    #rar[[i-2+10]] <- x2
    #str(x2)
    time <- c(1:i)
    for(j in 1:i){
        time[j] <- (as.numeric(x)[j] + as.numeric(x)[j+1])/2
    }
    xx2 <- as.data.frame(x2)[1,]
    se <- as.data.frame(x2)[2,]
    #if(i == 3){
        #rarecurve(comm2, label = TRUE, step = 20, cex = 0.1, las = 2)
        #abline(v = 25)}
    m <- lm(t(xx2)~ time)
    summary(m)
    corr <- cor(t(xx2), time, use = "complete.obs")
    cor_dis <- c(1:1000)
    for (k in 1:1000){
        cor_dis[k] <- cor(sample(t(xx2),i), time, use = "complete.obs")
    }
    #hist(cor_dis)
    #lines(c(corr, corr), c(0,200), col = "red")
    (p <- pnorm(corr, mean = mean(cor_dis), sd = sd(cor_dis)))
    #with lm
    #if(summary(m)$coef[8]> 0.05 | is.na(summary(m)$coef[8]) == TRUE){
    #abline(m, lty = 2)}
    #else{
    #abline(m, lty = 1)}    
    #with corrs
    #if(p > 0.05){
      # abline(m, lty = 2)}
   # else{
       # abline(m, lty = 1)}    
   # if(i == 10){
        #rarecurve(comm2, label = TRUE, step = 20, cex = 0.1, las = 2)
        #abline(v = 25)}    
    #rarecurve(comm2)
    #abline(v = 1000)
}
    plot(time,t(xx2), xlab = "", las = 2, ylab = "Number of exotic bee species", main = "", xaxt = "n", ylim = c(min(xx2, na.rm = TRUE)- max(se), max(xx2, na.rm = TRUE)+ max(se)))
    axis(1, at = time ,labels = lab, las = 2, cex.axis = 0.7)
    arrows(time,as.numeric(xx2)+as.numeric(se), time, as.numeric(xx2)-as.numeric(se), angle=90, length = 0)
    abline(m, lty = 2)

############################################
#resampling analysis for flies
############################################

@@

    plot(time,t(xx2), xlab = "", las = 2, ylab = "Number of native fly species", main = "", xaxt = "n", ylim = c(min(xx2, na.rm = TRUE)- max(se), max(xx2, na.rm = TRUE)+ max(se)))
    axis(1, at = time ,labels = lab, las = 2, cex.axis = 0.7)
    arrows(time,as.numeric(xx2)+as.numeric(se), time, as.numeric(xx2)-as.numeric(se), angle=90, length = 0)
    abline(m, lty = 2)

    ##########################################################################
    #Resampling Analysis for exotic flies
    ##########################################################################

@@

    plot(time,t(xx2), xlab = "", las = 2, ylab = "Number of exotic fly species", main = "", xaxt = "n", ylim = c(min(xx2, na.rm = TRUE)- max(se), max(xx2, na.rm = TRUE)+ max(se)))
    axis(1, at = time ,labels = lab, las = 2, cex.axis = 0.7)
    arrows(time,as.numeric(xx2)+as.numeric(se), time, as.numeric(xx2)-as.numeric(se), angle=90, length = 0)
    abline(m, lty = 2)

    ##########################################################################
    #END
    ##########################################################################

>>>>

---
title: "Historic collections as a tool for assessing the global pollination crisis"
author: "Bartomeus, I.^1^, Stavert, J.R.^2^, Ward, D.^2,3^, and Aguado, O.^4^"
csl: ../data/philb_style.xml
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
bibliography: ../data/references2.xml
---

@@

For illustrative purposes, we mapped the findings of these studies (Figure 1) to show the large differences in bee species richness worldwide, with bee diversity hotspots in mediterranean countries, against the paucity of countries for which we have any local or regional data on bee, syrphid or butterfly declines (raw data: Sup Mat 1). Except for outside of Europe and the US and for non-insect taxa, there are very few or no studies on pollinator declines using historical records. However, there are species-specific examples of historical losses from different parts of the world (e.g., *Bombus dalbhomi*; @Morales2013).
```{r, include=FALSE}
#update joinCountryData2Map to avoid printing results

joinCountryData2Map <- function (dF, joinCode = "ISO3", nameJoinColumn = "ISO3V10", 
    nameCountryColumn = "Country", suggestForFailedCodes = FALSE, 
    mapResolution = "coarse", projection = NA, verbose = FALSE) 
{
    mapWithData <- getMap(resolution = mapResolution)
    if (!is.na(projection)) 
        warning("the projection argument has been deprecated, returning Lat Lon, use spTransform from package rgdal as shown in help details or the FAQ")
    listJoinCodesNew <- c("ISO_A2", "ISO_A3", "FIPS_10_", "ADMIN", 
        "ISO_N3")
    listJoinCodesOld <- c("ISO2", "ISO3", "FIPS", "NAME", "UN")
    listJoinCodes <- c(listJoinCodesOld, listJoinCodesNew)
    if (joinCode %in% listJoinCodes == FALSE) {
        stop("your joinCode (", joinCode, ") in joinCountryData2Map() is not one of those supported. Options are :", 
            paste(listJoinCodes, ""), "\\n")
        return(FALSE)
    }
    joinCodeOld <- joinCode
    if (joinCode %in% listJoinCodesOld) {
        joinCode <- listJoinCodesNew[match(joinCode, listJoinCodesOld)]
    }
    if (is.na(match(nameJoinColumn, names(dF)))) {
        stop("your chosen nameJoinColumn :'", nameJoinColumn, 
            "' seems not to exist in your data, columns = ", 
            paste(names(dF), ""))
        return(FALSE)
    }
    dF[[joinCode]] <- as.character(dF[[nameJoinColumn]])
    dF[[joinCode]] <- gsub("[[:space:]]*$", "", dF[[joinCode]])
    if (joinCode == "ADMIN") {
        dF$ISO3 <- NA
        for (i in 1:nrow(dF)) dF$ISO3[i] = rwmGetISO3(dF[[joinCode]][i])
        joinCode = "ISO3"
        nameCountryColumn = nameJoinColumn
    }
    matchPosnsInLookup <- match(as.character(dF[[joinCode]]), 
        as.character(mapWithData@data[[joinCode]]))
    failedCodes <- dF[[joinCode]][is.na(matchPosnsInLookup)]
    numFailedCodes <- length(failedCodes)
    numMatchedCountries <- nrow(dF) - numFailedCodes
    #cat(numMatchedCountries, "codes from your data successfully matched countries in the map\\n")
    failedCountries <- dF[[nameCountryColumn]][is.na(matchPosnsInLookup)]
    failedCountries <- cbind(failedCodes, failedCountries = as.character(failedCountries))
    #cat(numFailedCodes, "codes from your data failed to match with a country code in the map\\n")
    if (verbose) 
        print(failedCountries)
    matchPosnsInUserData <- match(as.character(mapWithData@data[[joinCode]]), 
        as.character(dF[[joinCode]]))
    codesMissingFromUserData <- as.character(mapWithData@data[[joinCode]][is.na(matchPosnsInUserData)])
    countriesMissingFromUserData <- as.character(mapWithData@data[["NAME"]][is.na(matchPosnsInUserData)])
    numMissingCodes <- length(codesMissingFromUserData)
    #cat(numMissingCodes, "codes from the map weren't represented in your data\\n")
    mapWithData@data <- cbind(mapWithData@data, dF[matchPosnsInUserData, 
        ])
    invisible(mapWithData)
}
```

```{r map, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(rworldmap)
map <- read.csv("../data/bee_richness.csv", sep = ";")
size <- read.csv("../data/country_area.csv")
colnames(size) <- c("country", "size")
map0 <- merge(map, size, all.x = TRUE)
#remove some countries with doubtful sampling
#map0 <- subset(map0, !country %in% c("Kuwait", "Guinea-Bissau", "Chad", "Mali", "Mauritania"))
#map0[which(is.na(map0$size)),] #some can be rescued?
#scatter.smooth(log(map0$species) ~ log(map0$size+1))
m <- lm(log(map0$species) ~ log(map0$size+1), na.action = na.exclude) 
map0$res <- residuals(m)
#subset(map0, res < -2)
map2 <- joinCountryData2Map(map0, joinCode="NAME", nameJoinColumn = "country",
                            nameCountryColumn = "distribution", verbose = FALSE)
#need to use ISO's to make sure we don't carry over NA's due to country misspellings
#have plotted countries with missing values as grey (shows we have richness for pretty much everywhere)
par(mfrow = c(3,2), mar = c(2.1, 3.1, 3.1, 2.1))
map3 <- mapCountryData(map2, nameColumnToPlot = "res", missingCountryCol = "white",
               mapTitle = "Bee species richness", numCats = 25, catMethod = "pretty", addLegend = FALSE)
#add a modified legend using the same initial parameters as mapCountryData
do.call(addMapLegend, c(map3, legendLabels="all", legendWidth=0.5, labelFontSize=0.8, legendMar = 4))
mtext("a)", side = 3, adj = 0.05, 
    line = -1.3)
#need to add all the pannels.
library(plyr)
library(dplyr)
library(reshape2)

@@

map.bee <- joinCountryData2Map(declines.bee.2, joinCode="NAME", nameJoinColumn = "country",
                                   nameCountryColumn = "distribution", verbose = FALSE)

#average richness change
colorss <- c(heat.colors(9))
map.bee.ave <- mapCountryData(map.bee, nameColumnToPlot = "average", colourPalette= colorss,
                                  mapTitle = "Average bee richness change", catMethod = "pretty", addLegend = F)
do.call(addMapLegend, c(map.bee.ave, legendLabels="all", legendWidth=0.5, labelFontSize=0.8, legendMar = 4))
mtext("b)", side = 3, adj = 0.05, 
    line = -1.3)

#syrphid richness map
map_s <- subset(map0, is.na(syrphid_complete))
#scatter.smooth(log(map_s$syrphid) ~ log(map_s$size+1))
m <- lm(log(map_s$syrphid) ~ log(map_s$size+1), na.action = na.exclude) 
map_s$res <- residuals(m)
#subset(map0, res < -2)
map2 <- joinCountryData2Map(map_s, joinCode="NAME", nameJoinColumn = "country",
                            nameCountryColumn = "distribution", verbose = FALSE)
#need to use ISO's to make sure we don't carry over NA's due to country misspellings
#have plotted countries with missing values as grey (shows we have richness for pretty much everywhere)
map3 <- mapCountryData(map2, nameColumnToPlot = "res", missingCountryCol = "white",
               mapTitle = "Syrphid species richness", numCats = 25, catMethod = "pretty", addLegend = FALSE, mapRegion = "Europe")
#add a modified legend using the same initial parameters as mapCountryData
do.call(addMapLegend, c(map3, legendLabels="all", legendWidth=0.5, labelFontSize=0.8, legendMar = 4))
mtext("c)", side = 3, adj = 0.05, 
    line = -1.3)

@@

map.syrphid <- joinCountryData2Map(declines.syrphid.2, joinCode="NAME", nameJoinColumn = "country",
                            nameCountryColumn = "distribution", verbose = FALSE)

#average richness change
colorss <- c(heat.colors(5), c("#5e977c", "#3d88a5", "#1c79cf"))
map.syrphid.ave <- mapCountryData(map.syrphid, nameColumnToPlot = "average", colourPalette= colorss,
                       mapTitle = "Average syrphid richness change", catMethod = "pretty", addLegend = F, mapRegion = "Europe")
do.call(addMapLegend, c(map.syrphid.ave, legendLabels="all", legendWidth=0.5, labelFontSize=0.8, legendMar = 4))
mtext("d)", side = 3, adj = 0.05, 
    line = -1.3)

#butterfly richness
#scatter.smooth(log(map0$buterflies) ~ log(map0$size+1))
m <- lm(log(map0$syrphid) ~ log(map0$size+1), na.action = na.exclude) 
map0$res <- residuals(m)
#subset(map0, res < -2)
map2 <- joinCountryData2Map(map0, joinCode="NAME", nameJoinColumn = "country",
                            nameCountryColumn = "distribution", verbose = FALSE)
#need to use ISO's to make sure we don't carry over NA's due to country misspellings
#have plotted countries with missing values as grey (shows we have richness for pretty much everywhere)
map3 <- mapCountryData(map2, nameColumnToPlot = "res", missingCountryCol = "white",
               mapTitle = "Butterfly species richness", numCats = 25, catMethod = "pretty", addLegend = FALSE, mapRegion = "Europe")
#add a modified legend using the same initial parameters as mapCountryData
do.call(addMapLegend, c(map3, legendLabels="all", legendWidth=0.5, labelFontSize=0.8, legendMar = 4))
mtext("e)", side = 3, adj = 0.05, 
    line = -1.3)

@@

map.butterfly <- joinCountryData2Map(declines.butterfly.2, joinCode="NAME", nameJoinColumn = "country",
                                     nameCountryColumn = "distribution", verbose = FALSE)

#average richness change
colorss <- c(heat.colors(6))
map.butterfly.ave <- mapCountryData(map.butterfly, nameColumnToPlot = "average", colourPalette=colorss,
                                    mapTitle = "Average butterfly richness change", catMethod = "pretty", addLegend = F, mapRegion = "Europe")
do.call(addMapLegend, c(map.butterfly.ave, legendLabels="all", legendWidth=0.5, labelFontSize=0.8, legendMar = 4))
mtext("f)", side = 3, adj = 0.05, 
    line = -1.3)
par(mfrow = c(1,1))
```

**Figure 1:** Global map showing a), c) and e) bee, syrphid and butterfly species richness per area (Data from http://www.discoverlife.org, IUCN and http://www.syrphidae.com/). ALl richness levels are calculated as the residuals of the log-log regression between bee species richness per country and country size. This correction accounts for the species-area relationship. Warmer colours indicate higher bee diversity. Note that some African countries may have incomplete listed faunas and that Alaska is included with US values. Countries with available historical changes in (b) bee, (d) syrphid and (f) butterfly richness within the last 100 years. Warmer colours indicate steeper average declines. Countries without data are coloured in white.

@@

#table?
#lhf[which(log(lhf$post1980) > 7.5 & log(lhf$pre1980) < 7.5),]
#lhf[which(log(lhf$post1980) > 7.5 & log(lhf$pre1980) > 7.5),]
#nrow(lhf[which(log(lhf$post1980) < 6.9 & log(lhf$pre1980) < 6.9),])
# poland switcherland need old records! > 10000 vs 300
# sri lanka and french guayana, 
# nicaragua, venezuela, Ecuador..>2000 vs >500.
# Colombia & Panama, zimbawe > 3000 vs >1500
#lhf[which(lhf$country == "ES" | lhf$country == "NZ"),]
col <- ifelse(lhf$country == "ES" | lhf$country == "NZ", "red", "black")
col <- ifelse(log(lhf$post1980) > 7.5 & log(lhf$pre1980) > 7.5, col, "grey")
pchs <- ifelse(lhf$country == "ES" | lhf$country == "NZ", 19, 1)
#par(mfrow = c(2,2))
#plot(log(lhf$post1980+1), log(lhf$pre1980+1), xlab = "post1980", ylab = "pre1980", col = col, pch = pchs, las = 1)
#temp <- subset(lhf, country %in% c("ES", "NZ"))
#text(x = temp$log_post1980, y = temp$log_pre1980+2 , labels = temp$country, col = "red")
#abline(h = 7.5, v = 7.5)
#36 countries!
#exp(7.5) #> 1800 records in both axes!
```

@@

#plot all
par(mfrow = c(2,2),  mar = c(2.1, 3.1, 3.1, 2.1))
plot(log(lhf$post1980+1), log(lhf$pre1980+1), xlab = "post1980", ylab = "pre1980", col = col, pch = pchs, las = 1)
temp <- subset(lhf, country %in% c("ES", "NZ"))
text(x = temp$log_post1980, y = temp$log_pre1980+2 , labels = temp$country, col = "red")
abline(h = 7.5, v = 7.5)
mtext("a)", side = 3, adj = 0.05, 
    line = -1.3)
boxplot(log(data2$pre1980), log(data2$post1980), names = c("pre1980", "post1980"), ylab = "log rarefied species richness", 
            las = 1, border = "light grey")
points(x = c(rep(1,nrow(data2)), rep (2,nrow(data2))), y = c(log(data2$pre1980), log(data2$post1980)), pch = 1)
for(i in 1:nrow(data2)){
    lines(x = 1:2, y = c(log(data2$pre1980[i]), log(data2$post1980[i])), 
          col = ifelse((data2$pre1980[i] - data2$post1980[i]) > 1, "red", "blue"))
}
mtext("b)", side = 3, adj = 0.05, 
    line = -1.3)
#map
data2$change <- ((data2$post1980 - data2$pre1980)*100 / data2$pre1980)
#mean(data2$change)
map.gbif <- joinCountryData2Map(data2, joinCode="ISO2", nameJoinColumn = "countries",
                            nameCountryColumn = "countries", verbose = FALSE)

#average richness change
#colorss <- c("firebrick", "red", "orangered", "orange", "yellow", "wheat", "green", "blue", "darkblue")
colorss <- c(heat.colors(5), c("#5e977c", "#3d88a5", "#1c79cf"))
map.gbif.plot <- mapCountryData(map.gbif, nameColumnToPlot = "change", colourPalette= colorss,
                       mapTitle = "Average bee richness change (%)", catMethod = "pretty", addLegend = F)
do.call(addMapLegend, c(map.gbif.plot, legendLabels="all", legendWidth=0.5))
mtext("c)", side = 3, adj = 0.05, 
    line = -1.3)
par(mfrow = c(1,1))

@@

In contrast to Spain, New Zealand is an isolated oceanic archipelago, with a distinctive pollinator biota and a unique history of human occupation. Much of New Zealand’s pollinator fauna is also relatively depauperate. For example, New Zealand has only 28 native bee species [@Donovan2007], which is a fraction of nearby Australia’s c. 1600 species [@Ascher2018]. However, New Zealand has a surprisingly high diversity of flies (Diptera), which are important pollinators in many ecosystems [@Newstrom2005]. Thus, New Zealand provides a unique system to study long-term changes in pollinator communities, and is unlike continental Europe and the US, which have been the focus of an overwhelming majority of pollinator decline studies.

@@

}
    plot(time,t(xx2), xlab = "", las = 2, ylab = "Number of native bee species", main = "", xaxt = "n", ylim = c(min(xx2, na.rm = TRUE)- max(se), max(xx2, na.rm = TRUE)+ max(se)))
    axis(1, at = time ,labels = lab, las = 2, cex.axis = 0.7)
    arrows(time,as.numeric(xx2)+as.numeric(se), time, as.numeric(xx2)-as.numeric(se), angle=90, length = 0)
    abline(m, lty = 2)
mtext("a)", side = 3, adj = 0.05, 
    line = -1.3)
#############################################
#Resampling Analysis for exotic species 
#############################################

#length(which(exotic$year < 1900))
exotic <- subset(exotic, year > 1800)
for (i in 9){ #error saying break are not unique - how to fix?
    break1 <- quantile(exotic$year, probs = seq(0,1,1/6), na.rm = TRUE)
    x <- c(1916,1952,1953,1969,1971,1975,1978,1983,1991,2011)
    lab <- c(1:i)
    for(j in 1:i){
        lab[j] <- paste(as.numeric(x)[j],as.numeric(x)[j+1], sep = "-")
    }
    exotic$time_period_quant  <- cut(exotic$year, breaks = x, labels = lab)
    comm2 <- table(exotic$time_period_quant, factor(exotic$Gen_sp))
    (x2 <- rarefy(comm2,25, se = TRUE)) #same result as my resampling...
    #rar[[i-2+10]] <- x2
    #str(x2)
    time <- c(1:i)
    for(j in 1:i){
        time[j] <- (as.numeric(x)[j] + as.numeric(x)[j+1])/2
    }
    xx2 <- as.data.frame(x2)[1,]
    se <- as.data.frame(x2)[2,]
    #if(i == 3){
        #rarecurve(comm2, label = TRUE, step = 20, cex = 0.1, las = 2)
        #abline(v = 25)}
    m <- lm(t(xx2)~ time)
    summary(m)
    corr <- cor(t(xx2), time, use = "complete.obs")
    cor_dis <- c(1:1000)
    for (k in 1:1000){
        cor_dis[k] <- cor(sample(t(xx2),i), time, use = "complete.obs")
    }
    #hist(cor_dis)
    #lines(c(corr, corr), c(0,200), col = "red")
    (p <- pnorm(corr, mean = mean(cor_dis), sd = sd(cor_dis)))
    #with lm
    #if(summary(m)$coef[8]> 0.05 | is.na(summary(m)$coef[8]) == TRUE){
    #abline(m, lty = 2)}
    #else{
    #abline(m, lty = 1)}    
    #with corrs
    #if(p > 0.05){
      # abline(m, lty = 2)}
   # else{
       # abline(m, lty = 1)}    
   # if(i == 10){
        #rarecurve(comm2, label = TRUE, step = 20, cex = 0.1, las = 2)
        #abline(v = 25)}    
    #rarecurve(comm2)
    #abline(v = 1000)
}
    plot(time,t(xx2), xlab = "", las = 2, ylab = "Number of exotic bee species", main = "", xaxt = "n", ylim = c(min(xx2, na.rm = TRUE)- max(se), max(xx2, na.rm = TRUE)+ max(se)))
    axis(1, at = time ,labels = lab, las = 2, cex.axis = 0.7)
    arrows(time,as.numeric(xx2)+as.numeric(se), time, as.numeric(xx2)-as.numeric(se), angle=90, length = 0)
    abline(m, lty = 2)
mtext("b)", side = 3, adj = 0.05, 
    line = -1.3)
############################################
#resampling analysis for flies
############################################

@@

    plot(time,t(xx2), xlab = "", las = 2, ylab = "Number of native fly species", main = "", xaxt = "n", ylim = c(min(xx2, na.rm = TRUE)- max(se), max(xx2, na.rm = TRUE)+ max(se)))
    axis(1, at = time ,labels = lab, las = 2, cex.axis = 0.7)
    arrows(time,as.numeric(xx2)+as.numeric(se), time, as.numeric(xx2)-as.numeric(se), angle=90, length = 0)
    abline(m, lty = 2)
mtext("c)", side = 3, adj = 0.05, 
    line = -1.3)    
    ##########################################################################
    #Resampling Analysis for exotic flies
    ##########################################################################

@@

    plot(time,t(xx2), xlab = "", las = 2, ylab = "Number of exotic fly species", main = "", xaxt = "n", ylim = c(min(xx2, na.rm = TRUE)- max(se), max(xx2, na.rm = TRUE)+ max(se)))
    axis(1, at = time ,labels = lab, las = 2, cex.axis = 0.7)
    arrows(time,as.numeric(xx2)+as.numeric(se), time, as.numeric(xx2)-as.numeric(se), angle=90, length = 0)
    abline(m, lty = 2)
mtext("d)", side = 3, adj = 0.05, 
    line = -1.3)    
    ##########################################################################
    #END
    ##########################################################################
